---
title: "Approximation of Monte Carlo Method"
subtitle: "5361 Homework 2"
author: "Qinxiao Shi"
date: "09/04/2018"
abstract: "Consider approximation of standard normal distribution by the Monte Carlo methods."

output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1.2

## Functions

I'm going to compare standard Normal distribution:
$$\Phi(t)=\int_{-\infty}^{t} \frac{1}{\sqrt{2\pi}} e^{-y^{2}/2}dy$$
with the Monte Carlo methods, where $X_{i}$'s are iid $N(0,1)$ variables:
$$\hat{\Phi}(t)=\frac{1}{n} \sum_{i=1}^{n} I(X_{i}\leq t)$$
, when $n$ is large and goes larger.

Experiment with the approximation at $n\in{\lbrace 10^2, 10^3, 10^4 \rbrace}$ at $t\in{\lbrace0.0, 0.67, 0.84, 1.28, 1.65, 2.32, 2.58, 3.09, 3.72\rbrace}$



```{r table, echo=TRUE}
library(knitr)
library(kableExtra)

t <- c(0.0, 0.67, 0.84, 1.28, 1.65, 2.32, 2.58, 3.09, 3.72)
a <- array(data=NA, dim=length(t))

for (i in 1:length(t)){
a[i] <- pnorm(t[i], mean=0, sd=1)
}
t <- c("t=", t)
a <- c("F(x)=", a)

table1 <- array(c(t, a), dim = c(10, 2))

set.seed(20180913)
norm100 <- rnorm(100)
norm1000 <- rnorm(1000)
norm10000 <- rnorm(10000)

t <- c(0.0, 0.67, 0.84, 1.28, 1.65, 2.32, 2.58, 3.09, 3.72)
b <- array(data=NA, dim=length(t))
c <- array(data=NA, dim=length(t))
d <- array(data=NA, dim=length(t))

for (i in 1:length(t)){
b[i] <- sum(norm100 <= t[i])/100
c[i] <- sum(norm1000 <= t[i])/1000
d[i] <- sum(norm10000 <= t[i])/10000
}
t <- c("t=", t)
b <- c("Sum= (n=100)", b)
c <- c("Sum= (n=1000)", c)
d <- c("Sum= (n=10000)", d)

table2 <- array(c(t, b, c, d), dim = c(10, 4))

kable(table1, caption = 'The CDF of Standard Normal Distribution', booktabs=T) %>%
  row_spec(1, bold = T, hline_after = T)

kable(table2, caption = 'The Summation of Monte Carlo Methods', booktabs=T) %>%
  row_spec(1, bold = T, hline_after = T)
  
```


## Figures

It is visual from graphs of two equations:


```{r histgm, echo=TRUE}
hist(norm100, col = 'darkgray', border = 'white', main = "Histogram of N(0,1) When n=100")

hist(norm1000, col = 'darkgray', border = 'white', main = "Histogram of N(0,1) When n=1000")

hist(norm10000, col = 'darkgray', border = 'white', main = "Histogram of N(0,1) When n=10000")
```

Which is obvious that when n increasing, the distribution of samples random chose from standard normal distribution is more close to standard normal distribution, which is a continues distribution.

```{r boxplot, echo=TRUE}
group <- array(NA, dim=11100)
for (x in 1:100){
  group[x] <- "n = 100"
}
for (x in 101:1100){
  group[x] <- "n = 1000"
}
for (x in 1101:11100){
  group[x] <- "n = 10000"
}

norm <- c(norm100, norm1000, norm10000)
boxplot(norm~group, data = NULL,
        main = "Box Plot of N(0,1) with Difference Sample Size",
        xlab = "Sample Size",
        ylab = "Standard Normal Distribution")
```

The box plots show that when $n$ goes larger and larger, the distribution approaches a standard normal distribution. So the mean is more and more close to $0$, and distribution is symmetric on $0$.

Since the highly similarity of two distributions, cdf of $N(0,1)$ can be simulated by Monte Carlo method in calculation.

# Exercise 1.3

The double-precision floating-point format is is a computer number format, usually occupying 64 bits in computer memory.
$$(-1)^{sign}(1+\sum_{i=1}^{52}b_{52-i}2^{-i}) \times 2^{e-1023}$$

## .Machine$double.xmax 
Which is the largest normalized floating-point number.
In decimal system, it can be denoted as $$(-1)^0(1+(1-2^{-52}))2^{1023}$$

In binary system, it can be denoted as $0 \ 11111111110 \ 1111111111111111111111111111111111111111111111111111_2$

```{r xmax, echo=TRUE}
.Machine$double.xmax
```


## .Machine$double.xmin
Which is the smallest  non-vanishing normalized floating-point power of the radix.
In decimal system, it can be denoted as $$(-1)^0(1+1)2^{-1023}$$

In binary system, it can be denoted as $0 \ 00000000001 \ 0000000000000000000000000000000000000000000000000000_2$
```{r xmin, echo=TRUE}
.Machine$double.xmin

```


## .Machine$double.eps
Which is the smallest positive floating-point number $x$ such that $1+x \ !=1$.
In decimal system, it can be denoted as $$2^{-52}$$

In binary system, it can be denoted as $00000000001_2$
```{r eps, echo=TRUE}
.Machine$double.eps

```


## .Machine$double.neg.eps
Which is a small positive floating-point number $x$ such that $1-x \ !=1$.
In decimal system, it can be denoted as $$2^{-53}$$

In binary system, it can be denoted as $00000000000_2$
```{r negeps, echo=TRUE}
.Machine$double.neg.eps
```